---
title: 'STA4173: Project 2'
author: 'Your Name Here'
date: 'Last updated: `r Sys.Date()`'
output:
  html_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)

# pull in packages we need
library(gsheet)
library(tidyverse)
library(car)
library(ggpubr)

# you may need to install some packages first
# use the install.packages() function:
# e.g., install.packages("googlesheets4")

# NOTE -- you may need additional packages, depending on what is asked of you below :-)

options(scipen = 999)

# call in data
data <- as_tibble(gsheet2tbl("https://docs.google.com/spreadsheets/d/14rl-_jmNSoA67CPWcqQFku6GRDKDTBosAHLHok4PQCI/edit#gid=0"))

# count the number of categories the movie can be classified as
data$No_Cat <- str_count(data$Genre, ",")+1

# turn that into a factor variable
data$No_Cat <- as_factor(data$No_Cat)

# turn movie runtime into a numeric variable
data$Runtime_Num <- as.numeric(str_remove(data$Runtime, " min"))

# categorize movie runtime into < 2 hours, 2-3 hours, and > 3 hours
data$Hours <- if_else(data$Runtime_Num < 120, "< 120 min", 
                      if_else(data$Runtime_Num <= 180, "120-180 min", "> 180 min"))
```

#### 1. Describe the data using the following tables: 
```{r}
summaries_genre <- data %>% group_by(No_Cat) %>% 
  summarize(mean = mean(IMDB_Rating), sd = sd(IMDB_Rating), median = median(IMDB_Rating), IQR = IQR(IMDB_Rating))

summaries_hours <- data %>% group_by(Hours) %>% 
  summarize(mean = mean(IMDB_Rating), sd = sd(IMDB_Rating), median = median(IMDB_Rating), IQR = IQR(IMDB_Rating))
```
  
|                           | 1 Category | 2 Categories | 3 Categories |
|---------------------------|------------|--------------|--------------|
| Mean (Standard Deviation) | `r round(summaries_genre$mean[1],2)` (`r round(summaries_genre$sd[1],2)`)            |  INSERT 2 CAT. HERE            | INSERT 3 CAT. HERE             |
| Median (IQR)              | `r round(summaries_genre$median[1],2)` (`r round(summaries_genre$IQR[1],2)`)            |  INSERT 2 CAT. HERE            | INSERT 3 CAT. HERE             |

|                           | Less than 2 Hours | 2 to 3 Hours      | 3 or More Hours |
|---------------------------|-------------------|-------------------|-----------------|
| Mean (Standard Deviation) | `r round(summaries_hours$mean[1],2)` (`r round(summaries_hours$sd[1],2)`)                   | INSERT 2-3 HOURS HERE                  | INSERT 3+ HOURS HERE                 |
| Median (IQR)              |  `r round(summaries_hours$median[1],2)` (`r round(summaries_hours$IQR[1],2)`)                  | INSERT 2-3 HOURS HERE                  | INSERT 3+ HOURS HERE                 |

####  2. Construct side-by-side boxplots comparing the distribution for IMDB rating (IMDB_Rating) between the number of genres the movie is considered to be in (No_Cat) and, separately, the length of the movie (Hours) There is code for the number of categories available in the starter code; copy, paste, and edit to create the plots for the runtime.
```{r}
ggplot(data=data, aes(y= IMDB_Rating, x=No_Cat, fill=No_Cat)) +
  geom_boxplot() +
  theme_minimal() +
  labs(y = "IMDB Rating", x = "Number of Categories") 

# You will need to copy, paste, and edit to create the graph for the length of the movie
```

  
####  3. Describe the distributions. Are they skewed? Do you think there's a difference between the ratings of movies that fall into 1, 2, or 3 genres? What about the ratings of movies that are < 120 minutes as compared to those that are 120 minutes or longer? 
  
####  4. Use one-way ANOVA to determine if there's a difference in the average IMDB ratings (IMDB_Rating) between the number of genres a movie belongs to (No_Cat). Test at the &alpha; = 0.05 level. Include the formal test results (including the formal hypotheses), but in your interpretation, make sure that you explain the results such that a non-statistician can understand. 
  
####  5. Use the Kruskal-Wallis test to determine if there's a difference in median IMDB ratings (IMDB_Rating) between the number of genres a movie belongs to (No_Cat). Test at the &alpha; = 0.05 level. Include the formal test results (including the formal hypotheses), but in your interpretation, make sure that you explain the results such that a non-statistician can understand. 
  
#### 6. Construct the graph for assessing ANOVA assumptions. Are any of the ANOVA assumptions broken? Justify your answer and include statements about the specific graphs.
  
####  7. Use the Brown-Forsythe-Leveine test to formally test the variance assumption. Test at the &alpha; = 0.05$ level. Include the formal test results (including the formal hypotheses), but in your interpretation, make sure that you explain the results such that a non-statistician can understand. Include whether or not the assumption is broken.
  
####  8. Based on your answers to 6 and 7, which test should be reported? Explain why you're choosing the test you're choosing. (Hint: you will either choose ANOVA or Kruskal-Wallis.)
  
####  9. Use two-way ANOVA to show that there is not an interaction between number of genres (No_Cat) and length of movie (Hours). Test at the &alpha; = 0.05 level. Include the formal test results (including the formal hypotheses), but in your interpretation, make sure that you explain the results such that a non-statistician can understand.
  
####  10. Remove the interaction term from your ANOVA model. Are there main effects of number of genres (No_Cat) and length of movie (Hours)? Test at the &alpha; = 0.05 level. Include the formal test results (including the formal hypotheses), but in your interpretation, make sure that you explain the results such that a non-statistician can understand.
  
####  11. Perform posthoc testing on the significant main effect(s) by constructing the Tukey's graph of simultaneous intervals. In your interpretation, explain which groups are significantly different.
  
#### 12. Construct the graph for assessing ANOVA assumptions. Are any of the ANOVA assumptions broken? Justify your answer and include statements about the specific graphs.




