---
title: "Tests for Proportions"
subtitle: "STA4173 Lecture 10, Summer 2023"
execute:
  echo: true
  message: false
  warning: false
format: 
  revealjs:
    theme: uwf
    self-contained: true
    slide-number: false
    footer: "[STA4173 - Biostatistics](https://samanthaseals.github.io/STA4173)"
    width: 1600
    height: 900
    df-print: paged
    html-math-method: katex
editor: source
---

## Introduction

- Before today, we have focused on continuous outcomes.

- Now we will focus on categorical (or qualitative) outcomes.

- Today, we will review how to test one and two sample proportions.

- We will estimate a proportion using $\hat{p}$, $$\hat{p} = \frac{x}{n}$$.
    
- We will estimate the difference between two proportions using $\hat{p}_1 - \hat{p}_2$, $$\hat{p_1}- \hat{p_2} = \frac{x_1}{n_1} - \frac{x_2}{n_2}$$

## One-Sample Proportions

<center><img src = "images/L17a.png"></center>

## One-Sample Proportions

- **(1--$\alpha$)100% CI for a population proportion, $p$** $$\hat{p} \pm z_{\alpha/2} \sqrt{\frac{\hat{p}(1-\hat{p})}{n}}$$

- To construct this interval, we require both:
    - $n\hat{p}(1-\hat{p}) \ge 10$ and 
    - $n \le 0.05N$
    
## One-Sample Proportions

- We will use either the [`binom.test()`](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/binom.test) function or the [`prop.test()`](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/prop.test) function.

- If we have $n \le 30$,

```{r, eval = FALSE}
binom.test(x = [number of successes], 
           n = [sample size], 
           p = [hypothesized proportion], 
           alternative = [alternative])
```

- If we have $n > 30$,

```{r, eval = FALSE}
prop.test(x = [number of successes], 
          n = [sample size], 
          p = [hypothesized proportion], 
          alternative = [alternative],
          correct=FALSE)
```

## One-Sample Proportions

- Humira is a medication used to treat rheumatoid arthritis (RA). 

- In clinical trials of Humira, 705 subjects diagnosed with RA were administered 40 mg of Humira every other week. 

- Of the 705 subjects, 66 reported nausea as a side effect. 

- It is known that the proportion of RA subjects in similar studies receiving a placebo who report nausea as a side effect is 0.08. 

- Does the sample evidence represent significant evidence that a higher proportion of subjects receiving Humira experience nausea as a side effect than those taking a placebo? 
- Use the $\alpha = 0.05$ level of significance.

- **What are the important pieces?**

## One-Sample Proportions

- Humira is a medication used to treat rheumatoid arthritis (RA). 

- In clinical trials of Humira, 705 subjects diagnosed with RA were administered 40 mg of Humira every other week. 

- Of the <font color="#db0b9d">705 subjects</font>, <font color="#db0b9d">66 reported nausea as a side effect</font>. 

- It is known that the proportion of RA subjects in <font color="#db0b9d">similar studies receiving a placebo who report nausea as a side effect is 0.08</font>. 

- Does the sample evidence represent significant evidence that a higher proportion of subjects receiving Humira experience nausea as a side effect than those taking a placebo? 
- Use the $\alpha = 0.05$ level of significance.

- **What are the important pieces?**
    
## One-Sample Proportions

- What is the point estimate, \boldmath $\hat{p}$?

- What is the 95% confidence interval for $p$?

- Are there a higher proportion of subjects taking Humira experiencing nausea as a side effect than those taking a placebo?

## One-Sample Proportions

- Of the <font color="#db0b9d">705 subjects</font>, <font color="#db0b9d">66 reported nausea as a side effect</font>. 
    
- It is known that the proportion of RA subjects in <font color="#db0b9d">similar studies receiving a placebo who report nausea as a side effect is 0.08</font>. 

```{r}
prop.test(x = 66, 
          n = 705,
          correct=FALSE)
```

- Thus, the point estimate is $\hat{p} = 0.094$ and the 95% CI for $p$ is $(0.07, 0.12)$.

## One-Sample Proportions

- Of the <font color="#db0b9d">705 subjects</font>, <font color="#db0b9d">66 reported nausea as a side effect</font>. 
    
- It is known that the proportion of RA subjects in <font color="#db0b9d">similar studies receiving a placebo who report nausea as a side effect is 0.08</font>. 

```{r}
prop.test(x = 66, 
          n = 705,
          p = 0.08,
          alternative = "greater",
          correct=FALSE)
```

## One-Sample Proportions

- **Hypotheses:**

    - $H_0: \ p \le 0.08$ 
    - $H_1: \ p > 0.08$

- **Test Statistic and *p*-Value**

    - $\chi^2 = 1.77$ or $z_0 = \sqrt{1.7761} = `r round(sqrt(1.7761),3)`$
    - $p= 0.091$
    
- **Rejection Region**

    - Reject $H_0$ if $p < \alpha$; $\alpha=0.05$.

- **Conclusion / Interpretation**

    - Fail to reject $H_0$. 
    
    - There is not sufficient evidence to suggest that the proportion of subjects taking Humira who experience nausea is greater than 0.08.

## One-Sample Proportions

- Which do you think is easier to raise -- a boy or a girl? 

- When asked this question in 1947, 24\% of all Americans said raising a girl was easier. 
- In June 2018, the Gallup Organization surveyed 1500 adult Americans, of which 408 felt it was easier to raise a girl. 

- Does this result suggest the proportion of adult Americans who believe it is easier to raise a girl has changed since 1947? 

- Test at the $\alpha=0.10$ level.

- **What are the important pieces?**

## One-Sample Proportions

- Which do you think is easier to raise -- a boy or a girl? 

- When asked <font color="#db0b9d">this question in 1947, 24\%</font> of all Americans said raising a girl was easier. 

- In June 2018, the Gallup Organization <font color="#db0b9d">surveyed 1500 adult Americans</font>, of which <font color="#db0b9d">408 felt it was easier</font> to raise a girl. 

- Does this result suggest the proportion of adult Americans who believe it is easier to raise a girl has changed since 1947? 

- Test at the $\alpha=0.10$ level.

- **What are the important pieces?**

## One-Sample Proportions

```{r}
prop.test(x = 408, 
          n = 1500,
          p = 0.24,
          correct=FALSE)
```

- Thus, the point estimate is $\hat{p} = 0.272$ and the 95% CI for $p$ is $(0.25, 0.30)$.

## One-Sample Proportions

- **Hypotheses**

    - $H_0: \ p = 0.24$ 
    - $H_1: \ p \ne 0.24$

- **Test Statistic and *p*-Value**

    - $\chi^2 = 8.42$ or $z_0 = \sqrt{8.4211} = `r round(sqrt(8.4211),3)`$
    - $p= 0.004$

- **Rejection Region**

    - Reject $H_0$ if $p < \alpha$; $\alpha=0.10$.

- **Conclusion / Interpretation**

    - Reject $H_0$. 
    
    - There is sufficient evidence to suggest that the proportion of adult Americans who believe that it is easier to raise a girl has changed since 1947.
    

## Two-Sample Proportions

<center><img src = "images/L17b.png"></center>

## Two-Sample Proportions

- **(1--$\alpha$)100% CI for $p_1-p_2$** $$(\hat{p}_1 - \hat{p}_2) \pm z_{\alpha/2} \sqrt{\frac{\hat{p}_1 (1-\hat{p}_1)}{n_1} + \frac{\hat{p}_2(1-\hat{p}_2)}{n_2}}$$ where

    - The point estimate, $\hat{p}_1-\hat{p}_2$, is computed as $$\hat{p}_1 - \hat{p}_2 = \frac{x_1}{n_1} - \frac{x_2}{n_2},$$ and

    - $x_i$ is the number of individuals in group $i$ with a specified characteristic
    
    - $n_i$ is the sample size for group $i$

- To construct this interval, we require: 

    - $n_1 \hat{p}_1(1-\hat{p}_1) \ge 10$ and 
    - $n_2 \hat{p}_2(1-\hat{p}_2) \ge 10$.

## Two-Sample Proportions

- We again use the [`prop.test()`](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/prop.test) function.

```{r, echo = TRUE, eval = FALSE}
prop.test(x = c(x_1, x_2), 
          n = c(n_1, n_2), 
          alternative = [alternative],
          correct=FALSE)
```

## Two-Sample Proportions

- In clinical trials of Nasonex, 3774 adult and adolescent allergy patients (patients 12 years and older) were randomly divided into two groups. 

    - The patients in group 1 (experimental group) received 200 $\mu$g of Nasonex.
    
        - Of the 2103 patients in the experimental group, 547 reported headaches as a side effect.
    
    - The patients in group 2 (control group) received a placebo.  
    
        - Of the 1671 patients in the control group, 368 reported headaches as a side effect. 
        
- Is there evidence to conclude that the proportion of Nasonex users who experienced headaches as a side effect is greater than the proportion in the control group?

- Test at the $\alpha= 0.05$ level of significance.

- **What are the important pieces?**

## Two-Sample Proportions

- In clinical trials of Nasonex, 3774 adult and adolescent allergy patients (patients 12 years and older) were randomly divided into two groups. 

    - The patients in <font color="#db0b9d">group 1</font> (experimental group) received 200 $\mu$g of Nasonex.
    
        - Of the <font color="#db0b9d">2103 patients</font> in the experimental group, <font color="#db0b9d">547 reported headaches as a side effect</font>.
    
    - The patients in <font color="#db0b9d">group 2</font> (control group) received a placebo.  
    
        - Of the <font color="#db0b9d">1671 patients</font> in the control group, <font color="#db0b9d">368 reported headaches as a side effect</font>. 
        
- Is there evidence to conclude that the proportion of Nasonex users who experienced headaches as a side effect is greater than the proportion in the control group?

- Test at the $\alpha= 0.05$ level of significance.

- **What are the important pieces?**

## Two-Sample Proportions

- Of the <font color="#db0b9d">2103 patients</font> in the experimental group, <font color="#db0b9d">547 reported headaches as a side effect</font>.

- Of the <font color="#db0b9d">1671 patients</font> in the control group, <font color="#db0b9d">368 reported headaches as a side effect</font>. 

```{r, echo = TRUE}
prop.test(x = c(547, 368), 
          n = c(2103, 1671), 
          correct=FALSE)
```
- Thus, $\hat{p}_{\text{Exp}} = 0.260$, $\hat{p}_{\text{Ctrl}} = 0.220$ and $\hat{p}_{\text{Exp}} - \hat{p}_{\text{Ctrl}} = `r round(0.2601046-0.2202274, 3)`$.

- The 95% CI for $\hat{p}_{\text{Exp}} - \hat{p}_{\text{Ctrl}}$ is $(0.013, 0.067)$.

## Two-Sample Proportions

```{r, echo = TRUE}
prop.test(x = c(547, 368), 
          n = c(2103, 1671), 
          alternative = "greater",
          correct=FALSE)
```

## Two-Sample Proportions

- **Hypotheses**

    - $H_0: \ p_{\text{Exp}} \le p_{\text{Ctrl}}$  
    - $H_1: \ p_{\text{Exp}} > p_{\text{Ctrl}}$ 

- **Test Statistic and *p*-value**

    - $\chi^2 = 8.06$ or $z_0 = \sqrt{8.0618} = `r round(sqrt(8.0618),3)`$
    - $p= 0.002$

- **Rejection Region**

    - Reject $H_0$ if $p < \alpha$; $\alpha=0.05$.

- **Conclusion / Interpretation**

    - Reject $H_0$. 
    
    - There is sufficient evidence to suggest that the proportion of Nasonex users who experienced headaches as a side effect is greater than that of the control group.

## Two-Sample Proportions

-  A professor from the Department of Art wants to determine citizen support of spending federal tax money on the arts. 

- Of a random sample of 220 women, 59 responded yes. 

- Another random sample of 175 men showed that 56 responded yes. 

- Does this information indicate a difference between the population proportion of women and the population proportion of men who favor spending more federal tax dollars on the arts? 

- Test at the $\alpha=0.01$ level. 

- **What are the important pieces?**

## Two-Sample Proportions

-  A professor from the Department of Art wants to determine citizen support of spending federal tax money on the arts. 

- Of a random sample of <font color="#db0b9d">220 women</font>, <font color="#db0b9d">59 responded yes</font>. 

- Another random sample of <font color="#db0b9d">175 men</font> showed that <font color="#db0b9d">56 responded yes</font>. 

- Does this information indicate a difference between the population proportion of women and the population proportion of men who favor spending more federal tax dollars on the arts? 

- Test at the $\alpha=0.01$ level. 

- **What are the important pieces?**
    
## Two-Sample Proportions

- Of a random sample of <font color="#db0b9d">220 women</font>, <font color="#db0b9d">59 responded yes</font>. 

- Another random sample of <font color="#db0b9d">175 men</font> showed that <font color="#db0b9d">56 responded yes</font>. 

```{r, echo = TRUE}
prop.test(x = c(59, 56), 
          n = c(220, 175), 
          correct=FALSE)
```

- Thus, $\hat{p}_{\text{W}} = 0.268$, $\hat{p}_{\text{M}} = 0.320$ and $\hat{p}_{\text{W}} - \hat{p}_{\text{M}} = `r round(0.2681818-0.3200000, 3)`$.

- The 95% CI for $\hat{p}_{\text{M}} - \hat{p}_{\text{W}}$ is $(-0.142, 0.039)$.
    
## Two-Sample Proportions

- **Hypotheses**

    - $H_0: \ p_{\text{W}} = p_{\text{M}}$  
    - $H_1: \ p_{\text{W}} \ne p_{\text{M}}$ 

- **Test Statistic and *p*-value**

    - $\chi^2 = 1.27$ or $z_0 = \sqrt{1.2681} = `r round(sqrt(1.2681),3)`$
    - $p= 0.260$

- **Rejection Region**

    - Reject $H_0$ if $p < \alpha$; $\alpha=0.01$.

- **Conclusion / Interpretation**

    - Fail to reject $H_0$. 
    
    - There is not sufficient evidence to suggest that there is a difference in the proportion of men and women who favor spending more federal tax dollars on the arts.

## Goodness-of-Fit

- The goodness-of-fit test allows us to determine if a frequency distribution follows a specific distribution.

    - This could be a named distribution (e.g., normal)

    - It could also be a distribution without a name (e.g., the probabilities are specified)

- Before we can perform the goodness-of-fit test, we must compute expected counts.$$E_i = n p_i$$

    - e.g., suppose that we expect 25\% of Skittles to be red; if we have 100 Skittles, we then expect 25 of them to be red.
    
## Goodness-of-Fit

- **Hypotheses**

    - $H_0:$ The random variable follows the specified distribution.
    - $H_1:$ The random variable does not follow the specified distribution.

- **Test Statistic**

    - $\chi^2_0 = \sum_{i=1}^k \frac{(O_i-E_i)^2}{E_i}$, where $O_i$ = observed and  $E_i$ = expected

- ***p*-Value**
    - $p = P[\chi^2_{k-1} \ge \chi^2_0]$, where $k$ = number of categories

- **Rejection Region**

    - Reject $H_0$ if $p < \alpha$

## Goodness-of-Fit

- In R, we will use the `chisq.test()` function and plug in both the counts and the expected probabilities

```{r, echo = TRUE, eval = FALSE}
counts <- c(O_1, O_2, ..., O_k) # create O_i vector
probs <- c(p_1, p_2, ..., p_k) # create p_i vector
chisq.test(counts, p = probs)
```

## Goodness-of-Fit

- Using the economy data, below (based on the 2017 Current Population Survey, adjusted for inflation), determine if there is evidence to suggest that the distribution of income has changed since 2000. 

- Test at the $\alpha = 0.05$ level of significance.

| Income | Observed | Probability |
|-|:-:|:-:|
| Under \$15,000 | 161 | 0.099 |
| \$15,000 - \$24,999 | 144 | 0.098 |
| \$25,000 - \$34,999 | 138 | 0.093 |
| \$35,000 - \$49,999 | 184 | 0.135 |
| \$50,000 - \$74,999 | 247 | 0.179 |
| \$75,000 - \$99,999 | 188 | 0.131 |
| \$100,000 - \$149,999 | 217 | 0.149 |
| \$150,000 - \$199,999 | 105 | 0.061 |
| Over \$200,000 | 116 | 0.055 |

## Goodness-of-Fit

```{r}
counts <- c(161, 144, 138, 184, 247, 188, 217, 105, 116) # create O_i vector
probs <- c(0.099, 0.098, 0.093, 0.135, 0.179, 0.131, 0.149, 0.061, 0.055) # create p_i vector
chisq.test(counts, p = probs)
```

## Goodness-of-Fit

- **Hypotheses**

    - $H_0:$ Distribution of income in 2017 follows the same distribution as 2000.
    - $H_1:$ Distribution of income in 2017 does not follow the same distribution as 2000.

- **Test Statistic and *p*-Value**

    - $\chi^2_0 = 20.693$ 
    - $p = 0.008$

- **Rejection Region**

    - Reject $H_0$ if $p < \alpha$; $\alpha=0.05$

- **Conclusion/Interpretation**

    - Reject $H_0$. 
    
    - There is sufficient evidence to suggest that the distribution of income in 2017 does not follow the same distribution as in 2000.
    
## Goodness-of-Fit

- An obstetrician wants to know whether the proportion of children born on each day of the week is the same. 

- She randomly selects 500 birth records and obtains the data shown in the table below (based on data obtained from Vital Statistics of the United States, 2016). 

- Is there reason to believe that the day on which a child is born does not occur with equal frequency at the $\alpha = 0.01$ level of significance?

| Sun | Mon | Tues | Weds | Thurs | Fri | Sat |
|:-:|:-:|:-:|:-:|:-:|:-:|:-:|
| 46 | 76 | 83 | 81 | 81 | 80 | 53 |

- What about expected probabilities?

## Goodness-of-Fit

```{r}
counts <- c(46, 76, 83, 81, 81, 80, 53) # create O_i vector
probs <- c(1/7, 1/7, 1/7, 1/7, 1/7, 1/7, 1/7) # create p_i vector
chisq.test(counts, p = probs)
```

## Goodness-of-Fit

- **Hypotheses**

    - $H_0: \ p_M = p_T = ... = p_{Su} = \frac{1}{7}$.
    - $H_1:$ At least one proportion is different.


- **Test Statistic and *p*-Value**

    - $\chi^2_0 = 19.568$
    - $p = 0.003$
    
- **Rejection Region**

  - Reject $H_0$ if $p < \alpha$; $\alpha=0.01$

- **Conclusion/Interpretation**

    - Reject $H_0$.
    - There is sufficient evidence to suggest that at least one proportion is different.
    
## Test for Independence

- Let us now discuss testing two categorical variables to determine if a relationship exists. 

- Take, for example, this data:
<center><img src = "images/L18a.png"></center>

- We will use the $\chi^2$ test for independence to determine if happiness depends on marital status. 
    
## Test for Independence

- Like in the goodness-of-fit test, we will first compute expected values.

<center><img src = "images/L18b.png" width = 1000></center>

- We find the expected values,$$E_{ij} = \frac{R_i C_j}{n}$$ where
    - $R_i$ is the total for row $i$,
    - $C_j$ is the total for column $j$,
    - $n$ is the total sample size

## Test for Independence

- **Hypotheses**

    - $H_0$: There is not a relationship between [var 1] and [var 2].
    - $H_1$: There is a relationship between [var 1] and [var 2].

- **Test Statistic**

    - $\chi_0^2 = \sum_{i=1}^k \frac{(O_i-E_i)^2}{E_i}$

- **p-Value**

    - $p = \text{P}[\chi^2_{(r-1)(c-1)} \ge \chi^2_0]$

- **Rejection Region**

    -  Reject $H_0$ if $p < \alpha$
    
## Test for Independence

- If given the contingency table, we can enter it in a `matrix()` (see example) and use the `chisq.test()` function.

```{r, echo = TRUE, eval = FALSE}
chisq.test([matrix name])
```

- If given raw data, we can use the [`CrossTable()`](https://www.rdocumentation.org/packages/gmodels/versions/2.18.1.1/topics/CrossTable) function in the [`gmodels`](https://www.rdocumentation.org/packages/gmodels/versions/2.18.1.1) package.

    - Note: this function replicates PROC FREQ from SAS :)

```{r, eval = FALSE}
CrossTable([dataset$row var], [dataset$col var], 
           prop.chisq= FALSE,  # turn off proportion contributed to chi-square statistic
           prop.t = FALSE, # turn off total proportions
           chisq = TRUE) # request chi-square test
```

## Test for Independence

```{r}
observed_table <- matrix(c(600, 63, 112, 144,
                           720, 142, 355, 459,
                           93, 51, 119, 127), 
                         nrow = 3, ncol = 4, byrow = T)
# I prefer to include breaks to make it look like the table given just for checking purposes
# make sure you edit the number of rows (nrow) and columns (ncol)!

rownames(observed_table) <- c("Very Happy", "Pretty Happy", "Not Too Happy") # name rows
colnames(observed_table) <- c("Married", "Widowed", "Divorced/Separated", "Never Married") # name cols

observed_table # print table to make sure it is what we want

chisq.test(observed_table) # chi-squared test for independence
```

## Test for Independence

- **Hypotheses**

    - $H_0:$ Happiness does not depend on marital status.
    - $H_1:$ Happiness depends on marital status.

- **Test Statistic and *p*-Value**

    - $\chi^2_0 = 224.116$
    - $p < 0.001$

- **Rejection Region**

    - Reject $H_0$ if $p < \alpha$; $\alpha=0.05$

- **Conclusion/Interpretation**

    - Reject $H_0$. 
    
    - There is sufficient evidence to suggest that happiness depends on marital status.
    
## Test for Independence 

- To see how the `CrossTable()` function works, let's explore the Palmer penguin dataset.

```{r}
library(gmodels)
penguins <- palmerpenguins::penguins 
CrossTable(penguins$species, penguins$sex, 
           prop.chisq= FALSE,  # turn off proportion contributed to chi-square statistic
           prop.t = FALSE, # turn off total proportions
           chisq = TRUE) # request chi-square test
```


## Test for Independence

- **Hypotheses**

    - $H_0:$ Biological sex of penguin depends on penguin species.
    - $H_1:$ Biological sex of penguin does not depend on penguin species.

- **Test Statistic and *p*-Value**

    - $\chi^2_0 = 0.049$
    - $p = 0.976$

- **Rejection Region**

    - Reject $H_0$ if $p < \alpha$; $\alpha=0.05$

- **Conclusion/Interpretation**

    - Fail to reject $H_0$. 
    
    - There is not sufficient evidence to suggest that the biological sex of penguins depends on species of penguin.

## Two Dependent Proportions

- We previously discussed comparing two proportions that are independent.

- Let us now consider two dependent proportions.

    - This requires matched pairs.

- Example

    - We want to determine whether there is a difference between two ointments meant to treat poison ivy. 
    
    - Suppose we apply ointment A on one arm and ointment B on the other arm of each individual and record if the poison ivy cleared up. 
    
    - The individuals in each group are *not independent* because each ointment is applied to the same individual.

## Two Dependent Proportions 

- We will apply McNemar's test for two dependent proportions. 

- This requires a table as below:
<center><img src = "images/L18d.png"  width = 500></center>

- The data for the poison ivy example is as follows:
<center><img src = "images/L18e.png"  width = 700></center>

## Two Dependent Proportions

- **Hypotheses**

    - $H_0$: $p_1 = p_2$ 
    - $H_1$: $p_1 \ne p_2$

- **Test Statistic**

    - $\chi^2_0 = \frac{(f_{12} - f_{21})^2}{f_{12} + f_{21}}$
    
- ***p*-Value**

    - $p = P[\chi^2_{1} \ge \chi^2_0]$

- **Rejection Region**

    - Reject $H_0$ if $p < \alpha$

## Two Dependent Proportions

- We again take the matrix approach but now apply the [`mcnemar.test()`](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/mcnemar.test) function.

```{r, eval = FALSE}
mcnemar.test(observed_table, correct = FALSE)
```

## Two Dependent Proportions

- In our example data,

```{r}
observed_table <- matrix(c(293, 43,
                           31, 103), 
                         nrow = 2, ncol = 2, byrow = T)
# I prefer to include breaks to make it look like the table given just for checking purposes

rownames(observed_table) <- c("B Healed", "B Did Not Heal") # name rows
colnames(observed_table) <- c("A Healed", "A Did Not Heal") # name cols

observed_table # print table to make sure it is what we want

mcnemar.test(observed_table, correct = FALSE)
```

## Two Dependent Proportions

- **Hypotheses**

    - $H_0$: $p_{\text{A}} = p_{\text{B}}$ 
    - $H_1$: $p_{\text{A}} \ne p_{\text{B}}$

- **Test Statistic**

    - $\chi^2_0 = 1.9459$
    - $p = 0.976$

- **Rejection Region**

    - Reject $H_0$ if $p < \alpha$; $\alpha = 0.05$.

- **Conclusion/Interpretation**

    - Fail to reject $H_0$. 
    
    - There is not sufficient evidence to suggest that there is a difference between the two ointments.

## Two Dependent Proportions

- A recent General Social Survey asked the following two questions of a random sample of 1492 adult Americans under the hypothetical scenario that the government suspected that a terrorist act was about to happen: 

    - Do you believe that the authorities should have the right to tap people's telephone conversations?

    - Do you believe that the authorities should have the right to stop and search people on the street at random?

- The results are as follows, <center><img src = "images/L18f.png"></center>

- Do the proportion of people who agree with each scenario differ significantly? Use the $\alpha = 0.05$ level of significance.

## Two Dependent Proportions

```{r}
observed_table <- matrix(c(494, 335,
                           126, 537), 
                         nrow = 2, ncol = 2, byrow = T)

rownames(observed_table) <- c("Tap - Agree", "Tap - Disagree") # name rows
colnames(observed_table) <- c("Stop - Agree", "Stop - Disagree") # name cols

observed_table # print table to make sure it is what we want

mcnemar.test(observed_table, correct = FALSE) # McNemar's test
```

## Two Dependent Proportions

- **Hypotheses**

    - $H_0$: $p_{\text{Tap}} = p_{\text{Stop}}$ 
    - $H_1$: $p_{\text{Tap}} \ne p_{\text{Stop}}$

- **Test Statistic and *p*-Value**

    - $\chi^2_0 = 94.753$
    - $p <0.001$

- **Rejection Region**

    - Reject $H_0$ if $p < \alpha$; $\alpha = 0.05$.

- **Conclusion/Interpretation**

    - Reject $H_0$. 
    
    - There is  sufficient evidence to suggest that there is a difference in responses between the two questions.









