---
title: "Tests for Proportions: Part 1"
subtitle: "STA4173: Biostatistics"
execute:
  echo: true
  message: false
  warning: false
format: 
  revealjs:
    theme: uwf
    self-contained: true
    slide-number: false
    footer: "[STA4173 - Biostatistics](https://samanthaseals.github.io/STA4173)"
    width: 1600
    height: 900
    df-print: paged
    html-math-method: katex
editor: source
---

## Introduction

- Before today, we have focused on continuous outcomes.

- Now we will focus on categorical (or qualitative) outcomes.

- Today, we will review how to test one and two sample proportions.

- We will estimate a proportion using $\hat{p}$, $$\hat{p} = \frac{x}{n}$$.
    
- We will estimate the difference between two proportions using $\hat{p}_1 - \hat{p}_2$, $$\hat{p_1}- \hat{p_2} = \frac{x_1}{n_1} - \frac{x_2}{n_2}$$

## One-Sample Proportions

<center><img src = "images/L17a.png"></center>

## One-Sample Proportions

- **(1--$\alpha$)100% CI for a population proportion, $p$** $$\hat{p} \pm z_{\alpha/2} \sqrt{\frac{\hat{p}(1-\hat{p})}{n}}$$

- To construct this interval, we require both:
    - $n\hat{p}(1-\hat{p}) \ge 10$ and 
    - $n \le 0.05N$
    
## One-Sample Proportions

- We will use either the [`binom.test()`](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/binom.test) function or the [`prop.test()`](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/prop.test) function.

- If we have $n \le 30$,

```{r, eval = FALSE}
binom.test(x = [number of successes], 
           n = [sample size], 
           p = [hypothesized proportion], 
           alternative = [alternative])
```

- If we have $n > 30$,

```{r, eval = FALSE}
prop.test(x = [number of successes], 
          n = [sample size], 
          p = [hypothesized proportion], 
          alternative = [alternative],
          correct=FALSE)
```

## One-Sample Proportions

- Humira is a medication used to treat rheumatoid arthritis (RA). 

- In clinical trials of Humira, 705 subjects diagnosed with RA were administered 40 mg of Humira every other week. 

- Of the 705 subjects, 66 reported nausea as a side effect. 

- It is known that the proportion of RA subjects in similar studies receiving a placebo who report nausea as a side effect is 0.08. 

- Does the sample evidence represent significant evidence that a higher proportion of subjects receiving Humira experience nausea as a side effect than those taking a placebo? 
- Use the $\alpha = 0.05$ level of significance.

- **What are the important pieces?**

## One-Sample Proportions

- Humira is a medication used to treat rheumatoid arthritis (RA). 

- In clinical trials of Humira, 705 subjects diagnosed with RA were administered 40 mg of Humira every other week. 

- Of the <font color="#db0b9d">705 subjects</font>, <font color="#db0b9d">66 reported nausea as a side effect</font>. 

- It is known that the proportion of RA subjects in <font color="#db0b9d">similar studies receiving a placebo who report nausea as a side effect is 0.08</font>. 

- Does the sample evidence represent significant evidence that a higher proportion of subjects receiving Humira experience nausea as a side effect than those taking a placebo? 
- Use the $\alpha = 0.05$ level of significance.

- **What are the important pieces?**
    
## One-Sample Proportions

- What is the point estimate, \boldmath $\hat{p}$?

- What is the 95% confidence interval for $p$?

- Are there a higher proportion of subjects taking Humira experiencing nausea as a side effect than those taking a placebo?

## One-Sample Proportions

- Of the <font color="#db0b9d">705 subjects</font>, <font color="#db0b9d">66 reported nausea as a side effect</font>. 
    
- It is known that the proportion of RA subjects in <font color="#db0b9d">similar studies receiving a placebo who report nausea as a side effect is 0.08</font>. 

```{r}
prop.test(x = 66, 
          n = 705,
          correct=FALSE)
```

- Thus, the point estimate is $\hat{p} = 0.094$ and the 95% CI for $p$ is $(0.07, 0.12)$.

## One-Sample Proportions

- Of the <font color="#db0b9d">705 subjects</font>, <font color="#db0b9d">66 reported nausea as a side effect</font>. 
    
- It is known that the proportion of RA subjects in <font color="#db0b9d">similar studies receiving a placebo who report nausea as a side effect is 0.08</font>. 

```{r}
prop.test(x = 66, 
          n = 705,
          p = 0.08,
          alternative = "greater",
          correct=FALSE)
```

## One-Sample Proportions

- **Hypotheses:**

    - $H_0: \ p \le 0.08$ 
    - $H_1: \ p > 0.08$

- **Test Statistic and *p*-Value**

    - $\chi^2 = 1.77$ or $z_0 = \sqrt{1.7761} = `r round(sqrt(1.7761),3)`$
    - $p= 0.091$
    
- **Rejection Region**

    - Reject $H_0$ if $p < \alpha$; $\alpha=0.05$.

- **Conclusion / Interpretation**

    - Fail to reject $H_0$. 
    
    - There is not sufficient evidence to suggest that the proportion of subjects taking Humira who experience nausea is greater than 0.08.

## One-Sample Proportions

- Which do you think is easier to raise -- a boy or a girl? 

- When asked this question in 1947, 24\% of all Americans said raising a girl was easier. 
- In June 2018, the Gallup Organization surveyed 1500 adult Americans, of which 408 felt it was easier to raise a girl. 

- Does this result suggest the proportion of adult Americans who believe it is easier to raise a girl has changed since 1947? 

- Test at the $\alpha=0.10$ level.

- **What are the important pieces?**

## One-Sample Proportions

- Which do you think is easier to raise -- a boy or a girl? 

- When asked <font color="#db0b9d">this question in 1947, 24\%</font> of all Americans said raising a girl was easier. 

- In June 2018, the Gallup Organization <font color="#db0b9d">surveyed 1500 adult Americans</font>, of which <font color="#db0b9d">408 felt it was easier</font> to raise a girl. 

- Does this result suggest the proportion of adult Americans who believe it is easier to raise a girl has changed since 1947? 

- Test at the $\alpha=0.10$ level.

- **What are the important pieces?**

## One-Sample Proportions

```{r}
prop.test(x = 408, 
          n = 1500,
          p = 0.24,
          correct=FALSE)
```

- Thus, the point estimate is $\hat{p} = 0.272$ and the 95% CI for $p$ is $(0.25, 0.30)$.

## One-Sample Proportions

- **Hypotheses**

    - $H_0: \ p = 0.24$ 
    - $H_1: \ p \ne 0.24$

- **Test Statistic and *p*-Value**

    - $\chi^2 = 8.42$ or $z_0 = \sqrt{8.4211} = `r round(sqrt(8.4211),3)`$
    - $p= 0.004$

- **Rejection Region**

    - Reject $H_0$ if $p < \alpha$; $\alpha=0.10$.

- **Conclusion / Interpretation**

    - Reject $H_0$. 
    
    - There is sufficient evidence to suggest that the proportion of adult Americans who believe that it is easier to raise a girl has changed since 1947.
    

## Two-Sample Proportions

<center><img src = "images/L17b.png"></center>

## Two-Sample Proportions

- **(1--$\alpha$)100% CI for $p_1-p_2$** $$(\hat{p}_1 - \hat{p}_2) \pm z_{\alpha/2} \sqrt{\frac{\hat{p}_1 (1-\hat{p}_1)}{n_1} + \frac{\hat{p}_2(1-\hat{p}_2)}{n_2}}$$ where

    - The point estimate, $\hat{p}_1-\hat{p}_2$, is computed as $$\hat{p}_1 - \hat{p}_2 = \frac{x_1}{n_1} - \frac{x_2}{n_2},$$ and

    - $x_i$ is the number of individuals in group $i$ with a specified characteristic
    
    - $n_i$ is the sample size for group $i$

- To construct this interval, we require: 

    - $n_1 \hat{p}_1(1-\hat{p}_1) \ge 10$ and 
    - $n_2 \hat{p}_2(1-\hat{p}_2) \ge 10$.

## Two-Sample Proportions

- We again use the [`prop.test()`](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/prop.test) function.

```{r, echo = TRUE, eval = FALSE}
prop.test(x = c(x_1, x_2), 
          n = c(n_1, n_2), 
          alternative = [alternative],
          correct=FALSE)
```

## Two-Sample Proportions

- In clinical trials of Nasonex, 3774 adult and adolescent allergy patients (patients 12 years and older) were randomly divided into two groups. 

    - The patients in group 1 (experimental group) received 200 $\mu$g of Nasonex.
    
        - Of the 2103 patients in the experimental group, 547 reported headaches as a side effect.
    
    - The patients in group 2 (control group) received a placebo.  
    
        - Of the 1671 patients in the control group, 368 reported headaches as a side effect. 
        
- Is there evidence to conclude that the proportion of Nasonex users who experienced headaches as a side effect is greater than the proportion in the control group?

- Test at the $\alpha= 0.05$ level of significance.

- **What are the important pieces?**

## Two-Sample Proportions

- In clinical trials of Nasonex, 3774 adult and adolescent allergy patients (patients 12 years and older) were randomly divided into two groups. 

    - The patients in <font color="#db0b9d">group 1</font> (experimental group) received 200 $\mu$g of Nasonex.
    
        - Of the <font color="#db0b9d">2103 patients</font> in the experimental group, <font color="#db0b9d">547 reported headaches as a side effect</font>.
    
    - The patients in <font color="#db0b9d">group 2</font> (control group) received a placebo.  
    
        - Of the <font color="#db0b9d">1671 patients</font> in the control group, <font color="#db0b9d">368 reported headaches as a side effect</font>. 
        
- Is there evidence to conclude that the proportion of Nasonex users who experienced headaches as a side effect is greater than the proportion in the control group?

- Test at the $\alpha= 0.05$ level of significance.

- **What are the important pieces?**

## Two-Sample Proportions

- Of the <font color="#db0b9d">2103 patients</font> in the experimental group, <font color="#db0b9d">547 reported headaches as a side effect</font>.

- Of the <font color="#db0b9d">1671 patients</font> in the control group, <font color="#db0b9d">368 reported headaches as a side effect</font>. 

```{r, echo = TRUE}
prop.test(x = c(547, 368), 
          n = c(2103, 1671), 
          correct=FALSE)
```
- Thus, $\hat{p}_{\text{Exp}} = 0.260$, $\hat{p}_{\text{Ctrl}} = 0.220$ and $\hat{p}_{\text{Exp}} - \hat{p}_{\text{Ctrl}} = `r round(0.2601046-0.2202274, 3)`$.

- The 95% CI for $\hat{p}_{\text{Exp}} - \hat{p}_{\text{Ctrl}}$ is $(0.013, 0.067)$.

## Two-Sample Proportions

```{r, echo = TRUE}
prop.test(x = c(547, 368), 
          n = c(2103, 1671), 
          alternative = "greater",
          correct=FALSE)
```

## Two-Sample Proportions

- **Hypotheses**

    - $H_0: \ p_{\text{Exp}} \le p_{\text{Ctrl}}$  
    - $H_1: \ p_{\text{Exp}} > p_{\text{Ctrl}}$ 

- **Test Statistic and *p*-value**

    - $\chi^2 = 8.06$ or $z_0 = \sqrt{8.0618} = `r round(sqrt(8.0618),3)`$
    - $p= 0.002$

- **Rejection Region**

    - Reject $H_0$ if $p < \alpha$; $\alpha=0.05$.

- **Conclusion / Interpretation**

    - Reject $H_0$. 
    
    - There is sufficient evidence to suggest that the proportion of Nasonex users who experienced headaches as a side effect is greater than that of the control group.

## Two-Sample Proportions

-  A professor from the Department of Art wants to determine citizen support of spending federal tax money on the arts. 

- Of a random sample of 220 women, 59 responded yes. 

- Another random sample of 175 men showed that 56 responded yes. 

- Does this information indicate a difference between the population proportion of women and the population proportion of men who favor spending more federal tax dollars on the arts? 

- Test at the $\alpha=0.01$ level. 

- **What are the important pieces?**

## Two-Sample Proportions

-  A professor from the Department of Art wants to determine citizen support of spending federal tax money on the arts. 

- Of a random sample of <font color="#db0b9d">220 women</font>, <font color="#db0b9d">59 responded yes</font>. 

- Another random sample of <font color="#db0b9d">175 men</font> showed that <font color="#db0b9d">56 responded yes</font>. 

- Does this information indicate a difference between the population proportion of women and the population proportion of men who favor spending more federal tax dollars on the arts? 

- Test at the $\alpha=0.01$ level. 

- **What are the important pieces?**
    
## Two-Sample Proportions

- Of a random sample of <font color="#db0b9d">220 women</font>, <font color="#db0b9d">59 responded yes</font>. 

- Another random sample of <font color="#db0b9d">175 men</font> showed that <font color="#db0b9d">56 responded yes</font>. 

```{r, echo = TRUE}
prop.test(x = c(59, 56), 
          n = c(220, 175), 
          correct=FALSE)
```

- Thus, $\hat{p}_{\text{W}} = 0.268$, $\hat{p}_{\text{M}} = 0.320$ and $\hat{p}_{\text{W}} - \hat{p}_{\text{M}} = `r round(0.2681818-0.3200000, 3)`$.

- The 95% CI for $\hat{p}_{\text{M}} - \hat{p}_{\text{W}}$ is $(-0.142, 0.039)$.
    
## Two-Sample Proportions

- **Hypotheses**

    - $H_0: \ p_{\text{W}} = p_{\text{M}}$  
    - $H_1: \ p_{\text{W}} \ne p_{\text{M}}$ 

- **Test Statistic and *p*-value**

    - $\chi^2 = 1.27$ or $z_0 = \sqrt{1.2681} = `r round(sqrt(1.2681),3)`$
    - $p= 0.260$

- **Rejection Region**

    - Reject $H_0$ if $p < \alpha$; $\alpha=0.01$.

- **Conclusion / Interpretation**

    - Fail to reject $H_0$. 
    
    - There is not sufficient evidence to suggest that there is a difference in the proportion of men and women who favor spending more federal tax dollars on the arts.





