---
title: "Lab: Week 2"
author: "Your Name Here"
format: 
  html:
    self-contained: true
    message: false
    warning: false
    error: true
execute:
  echo: true
editor: source
---

Let's again consider the [Hollywood Age Gaps](https://github.com/rfordatascience/tidytuesday/blob/master/data/2023/2023-02-14/readme.md) dataset from [Tidy Tuesday](https://github.com/rfordatascience/tidytuesday). This dataset contains information on the age gap between movie love interests with the following rules:

- The two actors play characters with actual romantic interests (i.e., not just friends or coworkers).

- The youngest of the two actors is at least 17 years old.

- Animated characters are excluded.

Pulling in the data,

```{r} 
# do not change anything in this chunk
library(tidyverse)
age_gaps <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-02-14/age_gaps.csv') %>%
  mutate(both_sexes = case_when(character_1_gender == "woman" & character_2_gender == "man" ~ "woman & man",
                                character_1_gender == "man" & character_2_gender == "woman" ~ "woman & man",
                                character_1_gender == "woman" & character_2_gender == "woman" ~ "woman & woman",
                                character_1_gender == "man" & character_2_gender == "man" ~ "man & man"))
```

**1. (From last lab!) Find the mean and standard deviation of the age gap (*age_difference*) between romantic interests.**

```{r}

```

**2. Construct the 90% confidence interval for the population mean age gap (*age_difference*) Remember to state the resulting confidence interval.**

```{r}

```

<!-- 
Please fill in below.
Note that I will not give you a "fill in" after this lab.
I include it here to show you how I expect the answer to be given.
-->

The 90% CI for $\mu$ is (lower bound, upper bound).

**3. Construct the 95% confidence interval for the population mean age gap (*age_difference*) Remember to state the resulting confidence interval.**

```{r}

```

<!-- 
Please fill in below.
Note that I will not give you a "fill in" after this lab.
I include it here to show you how I expect the answer to be given.
-->

The 95% CI for $\mu$ is (lower bound, upper bound).

**4. Construct the 99% confidence interval for the population mean age gap (*age_difference*) Remember to state the resulting confidence interval.**

```{r}

```

<!-- 
Please fill in below.
Note that I will not give you a "fill in" after this lab.
I include it here to show you how I expect the answer to be given.
-->

The 99% CI for $\mu$ is (lower bound, upper bound).

**5. Use the appropriate hypothesis test to determine if the population age gap (*age_difference*) is greater than 10 years. Test at the $\alpha=0.10$ level. Remember to typeset your results.**

```{r}

```

<!-- 
Please fill in below.
Note that I will not give you a "fill in" after this lab.
I include it here to show you how I expect the answer to be given.
In the future, you can grab the code for this from slides.
(Remember that you have access to the .qmd files for the slides!
http://samanthaseals.github.io/STA4173)
-->

- **Hypotheses**

    - $H_0: \ \mu \text{ delete this and fill in the rest}$
    - $H_1: \ \mu \text{ delete this and fill in the rest}$
    
- **Test Statistic and *p*-Value**

    - $t_0 = \text{ delete this and insert the t value}$
    - $p = \text{ delete this and insert the p-value}$
    
- **Rejection Region**

    - Reject $H_0$ if $p < \alpha$; $\alpha = \text{ delete this and insert the $\alpha$ being used}$.
    
- **Conclusion/Interpretation**

    - [Fail to reject or Reject] $H_0$.
    
    - There [is or is not] sufficient evidence to suggest [alternative hypothesis in words].

**6. (Challenge!) Graph the confidence intervals found above. Hint: look at [this post on Stack Overflow](https://stackoverflow.com/questions/67846447/confidence-interval-using-ggplot2-manually) and adapt.**

```{r}

```














