---
title: "Lab: Week 3"
author: "Your Name Here"
format: 
  html:
    self-contained: true
execute:
  echo: true
editor: source
---

Monday's lecture: [dependent (paired) *t*-tests](https://github.com/samanthaseals/STA4173/blob/master/slides/L04.qmd)

Call in necessary libraries:

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
```

It is a commonly held belief that SUVs are safer than cars. If an SUV and car are in a collision, does the SUV sustain less damage (as suggested by the cost of repair)? The Insurance Institute for Highway Safety crashed SUVs into cars, with the SUV moving 10 miles per hour and the front of the SUV crashing into the rear of the car.

```{r}
suv <- c(1721, 1434,  850, 2329, 1415, 1470, 2884)
car <- c(1274, 2327, 3223, 2058, 3095, 3386, 4560)
data <- tibble(suv, car)
```

1. Explain why this is *dependent* data.

2. Describe the data using the group means and the mean of the difference. Remember to state the resulting point estimate (i.e., do not just have R output as your answer).

```{r}

```

3. Find the 95% confidence interval for the average difference in repair costs.

```{r}

```

4. Using the CI from part 3, is there sufficient evidence to suggest that the repair costs for cars is higher? Why or why not?

5. Use the paired $t$-test to determine if the repair costs for cars is higher. Test at the $\alpha=0.05$ level. Remember to:
- Check how the groups will be subtracted and state the alternative accordingly.
- Typeset the results of the hypothesis test (i.e., do not just have R output as your answer).

```{r}

```

6. Bonus! Construct a relevant data visualization.
