---
title: "Lab: Week 4"
author: "Your Name Here"
format: 
  html:
    self-contained: true
execute:
  echo: true
editor: source
---

Last Wednesday's lecture: [assessing assumptions for *t*-tests](https://github.com/samanthaseals/STA4173/blob/master/slides/L05.qmd)

Monday's lecture: [Wilcoxon rank sum / Mann-Whitney *U*](https://github.com/samanthaseals/STA4173/blob/master/slides/L06.qmd)

Wednesday's lecture: [Wilcoxon signed rank](https://github.com/samanthaseals/STA4173/blob/master/slides/L07.qmd)

Call in necessary libraries:

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
```

**Recall week 2's data (we used the independent two-sample *t*-test):**

- Ramp metering is a traffic engineering idea that requires cars entering a freeway to stop for a certain period of time before joining the traffic flow. The theory is that ramp metering controls the number of cars on the freeway and the number of cars accessing the freeway, resulting in a freer flow of cars, which ultimately results in faster travel times. To test whether ramp metering is effective in reducing travel times, engineers in Minneapolis, Minnesota, conducted an experiment in which a section of freeway had ramp meters installed on the on-ramps. The response variable for the study was speed of the vehicles. A random sample of 15 cars on the highway for a Monday at 6 p.m. with the ramp meters on and a second random sample of 15 cars on a different Monday at 6 p.m. with the meters off resulted in the following speeds (in miles per hour).

```{r}
speed <- c(28, 48, 56, 38, 31, 25, 43, 46, 50, 35, 55, 40, 42, 26, 47,
           24, 26, 42, 34, 37, 31, 47, 38, 17, 29, 23, 40, 37, 52, 41)
meter <- c(rep("on", 15), rep("off", 15))
data <- tibble(meter, speed)
```

1. Perform the appropriate nonparametric test to determine if the ramp meters increased speeds. Remember to:
- Check how the groups will be subtracted and state the alternative accordingly.
- Typeset the results of the hypothesis test (i.e., do not just have R output as your answer).

```{r}

```

2. Construct the appropriate q-q plot to assess the normality assumption.

```{r}

```

3. If you are reporting the results to the engineers, which test results are you going to report and why? (Hint: you will either report your *t*-test results from week 2 or you will report the nonparametric test from this week.) 

**Recall week 3's data (we used the paired *t*-test):**

- It is a commonly held belief that SUVs are safer than cars. If an SUV and car are in a collision, does the SUV sustain less damage (as suggested by the cost of repair)? The Insurance Institute for Highway Safety crashed SUVs into cars, with the SUV moving 10 miles per hour and the front of the SUV crashing into the rear of the car.

```{r}
suv <- c(1721, 1434,  850, 2329, 1415, 1470, 2884)
car <- c(1274, 2327, 3223, 2058, 3095, 3386, 4560)
data <- tibble(suv, car)
```

4. Perform the appropriate nonparametric test to determine if the cost of damage caused to cars is greater than that of SUVs. Remember to:
- Check how the groups will be subtracted and state the alternative accordingly.
- Typeset the results of the hypothesis test (i.e., do not just have R output as your answer).

```{r}

```

5. Construct the appropriate q-q plot to assess the normality assumption.

```{r}

```

6. If you are reporting the results to the Insurance Institute for Highway Safety, which test results are you going to report and why? (Hint: you will either report your *t*-test results from week 3 or you will report the nonparametric test from this week.) 

7. Bonus! Research how to construct a confidence interval for the median and implement it for the speed data. For full credit, link to where you got the code/information from. Remember to state the resulting CI (i.e., do not just have R output as your answer).

```{r}

```
