---
title: "Lab: Week 6"
author: "Your Name Here"
format: 
  html:
    self-contained: true
execute:
  echo: true
editor: source
---

Monday's lecture: [One-Way ANOVA](https://github.com/samanthaseals/STA4173/blob/master/slides/L08.qmd)

Wednesday's lecture: [Posthoc Testing](https://github.com/samanthaseals/STA4173/blob/master/slides/L09.qmd)

An obstetrician knew that there were more live births during the week than on weekends. She wanted to determine whether the mean number of births was the same for each of the five days of the week. She randomly selected eight dates for each of the five days of the week and obtained the following data:

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
births <- c(10456, 10023, 10691, 10283, 10265, 11189, 11198, 11465,
            11621, 11944, 11045, 12927, 12577, 11753, 12509, 13521,
            11084, 11570, 11346, 11875, 12193, 11593, 11216, 11818,
            11171, 11745, 12023, 12433, 12132, 11903, 11233, 12543,
            11545, 12321, 11749, 12192, 12422, 11627, 11624, 12543)
day <- c(rep("M", 8), rep("T", 8), rep("W", 8), rep("R", 8), rep("F", 8))
data <- tibble(day, births)
```

1. Find the group means and standard deviations. (Hint: look back at previous labs and projects for code to adapt.)

```{r}

```

2. Construct the one-way ANOVA table.

```{r}

```

3. Determine if the mean number of births is the same for each of the five days of the week. Test at the $\alpha=0.05$ level.

```{r}

```

4. Use Tukey's test to determine what pairwise differences exist. Test at the $\alpha=0.05$ level.

```{r}

```

5. Use Fisher's test to determine what pairwise differences exist. Test at the $\alpha=0.05$ level.

```{r}

```

6. What is the (theoretical) difference between Tukey's and Fisher's test?

7. Bonus! Construct a data visualization that will help you explain the analysis results.

```{r}

```






