---
title: "Lab: Week 7"
author: "Your Name Here"
format: 
  html:
    self-contained: true
execute:
  echo: true
editor: source
---

Monday's lecture: [ANOVA Assumptions](https://github.com/samanthaseals/STA4173/blob/master/slides/L10.qmd)

Wednesday's lecture: [Kruskal-Wallis](https://github.com/samanthaseals/STA4173/blob/master/slides/L11.qmd)

Recall last week's data -- an obstetrician knew that there were more live births during the week than on weekends. She wanted to determine whether the mean number of births was the same for each of the five days of the week. She randomly selected eight dates for each of the five days of the week and obtained the following data:

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
births <- c(10456, 10023, 10691, 10283, 10265, 11189, 11198, 11465,
            11621, 11944, 11045, 12927, 12577, 11753, 12509, 13521,
            11084, 11570, 11346, 11875, 12193, 11593, 11216, 11818,
            11171, 11745, 12023, 12433, 12132, 11903, 11233, 12543,
            11545, 12321, 11749, 12192, 12422, 11627, 11624, 12543)
day <- c(rep("M", 8), rep("T", 8), rep("W", 8), rep("R", 8), rep("F", 8))
data <- tibble(day, births)
```

1. State and explain the ANOVA assumptions.

2. Call in the `almost_sas()` function, shown in lecture, that will construct the graphs necessary to check the ANOVA assumptions.

```{r}

```

3. Use the `almost_sas()` function to create the graphs necessary to check the ANOVA assumptions.

```{r}

```

4. Perform the Brown-Forsyth-Levine test to determine if there is a difference between the variances. Test at the $\alpha=0.05$ level.

```{r}

```

5. Assess the assumptions using the graph constructed in question 3. Remember to comment on each graph, including which assumption is being assessed. (Note: no code is needed here - please just write your assessment.)

6. Using the results of your assessment in question 5, determine if we should report the ANOVA results from last week or if the Kruskal-Wallis is necessary. Make sure that you explain why your choice is the correct choice. (Note: no code is needed here.)

7. Use the Kruskal-Wallis test to determine if there is a difference in number of births between week days. Test at the $\alpha=0.05$ level.

```{r}

```

8. Use the Kruskal-Wallis posthoc procedure to determine what pairwise differences exist. Test at the $\alpha=0.05$ level.

```{r}

```

9. Bonus! Construct an appropriate data visualization to help with explaining the results from questions 7 and/or 8.

```{r}

```





