---
title: "Lab: June 29"
author: "Your Name Here"
format: 
  html:
    self-contained: true
execute:
  echo: true
  warning: false
  message: false
editor: source
---

Call in necessary libraries:

```{r}
library(tidyverse)
```

Ramp metering is a traffic engineering idea that requires cars entering a freeway to stop for a certain period of time before joining the traffic flow. The theory is that ramp metering controls the number of cars on the freeway and the number of cars accessing the freeway, resulting in a freer flow of cars, which ultimately results in faster travel times. To test whether ramp metering is effective in reducing travel times, engineers in Minneapolis, Minnesota, conducted an experiment in which a section of freeway had ramp meters installed on the on-ramps. The response variable for the study was speed of the vehicles. A random sample of 15 cars on the highway for a Monday at 6 p.m. with the ramp meters on and a second random sample of 15 cars on a different Monday at 6 p.m. with the meters off resulted in the following speeds (in miles per hour).

```{r}
speed <- c(48, 48, 56, 38, 31, 35, 43, 46, 50, 35, 55, 40, 42, 26, 47,
           34, 26, 42, 34, 37, 41, 47, 38, 17, 29, 23, 40, 37, 52, 41)
meter <- c(rep("on", 15), rep("off", 15))
data <- tibble(meter, speed)
head(data)
```

**1. Describe the data using the overall median and the group medians. Remember to state each of the medians (i.e., do not use R output as your only answer!).**

```{r}

```


**2. Perform the appropriate nonparametric test to determine if the ramp meters increased speeds. Test at the $\alpha=0.10$ level. Remember to typeset the results.**

```{r}

```

**3. Construct the appropriate q-q plot to assess the normality assumption for the *t*-test.**

```{r}

```

**4. If you are reporting the results to the engineers, which test results are you going to report and why? (Hint: you will either report your *t*-test results from the last lab or you will report the nonparametric test from this lab) **



It is a commonly held belief that SUVs are safer than cars. If an SUV and car are in a collision, does the SUV sustain less damage (as suggested by the cost of repair)? The Insurance Institute for Highway Safety crashed SUVs into cars, with the SUV moving 10 miles per hour and the front of the SUV crashing into the rear of the car.

```{r}
suv <- c(1721, 1434, 1850, 2329, 1415, 1470, 2884)
car <- c(1274, 2327, 3223, 2058, 3095, 3386, 4560)
data <- tibble(suv, car)
head(data)
```

**5. Describe the data using the overall median and the group medians. Remember to state each of the medians (i.e., do not use R output as your only answer!).**

```{r}

```

**6. Perform the appropriate nonparametric test to determine if the cost of damage caused to SUVs is less than that of cars. Test at the $\alpha=0.05$ level. Remember to typeset the results.**

```{r}

```

**7. Construct the appropriate q-q plot to assess the normality assumption for the *t*-test.**

```{r}

```

**8. If you are reporting the results to the Insurance Institute for Highway Safety, which test results are you going to report and why? (Hint: you will either report your *t*-test results from the last lab or you will report the nonparametric test from this lab.)**



**9. Bonus! Research how to construct a confidence interval for the median and implement it for the speed data. For full credit, link to where you got the code/information from. Remember to state the resulting CI (i.e., do not just have R output as your answer).**

```{r}

```


